The need for accuracy and smoothness in numerical simulations

### 2024-09-04 slides/ostrava/cckm-ppam-2024.tex

%% \frame{\frametitle{The secondary point of this talk}

%%   It might be possible to assert experimentally if
%%   \vspace{0.5cm}
%%   \begin{enumerate} \setlength\itemsep{1em}
%%   \item your functions are sufficiently smooth
%%   \item your equations are solved with sufficient accuracy
%%   \end{enumerate}
%% }

@@

Define 
\begin{equation}
  R_h := \frac{A_h - A_{2h}}{2^p - 1}, \quad F_h:= \frac{A_{2h} - A_{4h}}{A_h - A_{2h}} 
\end{equation}
If 
\begin{equation}
  E_h = T - A_h = \alpha h^p + \beta h^q + O(h^r), \quad h \rightarrow 0_+
\end{equation}
then
\begin{equation}
  \frac{E_h - R_h}{h^q} \rightarrow \text{constant}
\end{equation}
and
\begin{equation}
  \frac{2^p - F_h}{h^{q-p}} \rightarrow \text{constant}
\end{equation}
which implies
\begin{equation}
  \log | 2^p - F_h| \approx \log|\text{constant}| + (q-p) \log h
\end{equation}

@@

\frame{\frametitle{What can go wrong?}

  \begin{enumerate}\setlength\itemsep{1.5em}
  \item Proving the existence of an AEX is an exercise in
    \begin{center}
      Taylor expansion
    \end{center}
    We cannot complete the proof unless
    \begin{center}
      the functions are sufficiently smooth.
      \end{center}
    \item We cannot expect $\hat{A}_h$ to behave as predicted by the AEX unless
    $$ \hat{A}_h \approx A_h $$
    is a good approximation
    \end{enumerate}
}

@@

>>>>

%% \frame{\frametitle{The secondary point of this talk}
%%   It might be possible to assert experimentally if
%%   \vspace{0.5cm}
%%   \begin{enumerate} \setlength\itemsep{1em}
%%   \item your functions are sufficiently smooth
%%   \item your equations are solved with sufficient accuracy
%%   \end{enumerate}
%% }

@@

Define 
\begin{equation}
  R_h := \frac{A_h - A_{2h}}{2^p - 1}, \quad F_h:= \frac{A_{2h} - A_{4h}}{A_h - A_{2h}} 
\end{equation}
If 
\begin{equation}
  E_h = T - A_h = \alpha h^p + \beta h^q + O(h^r), \quad h \rightarrow 0_+
\end{equation}
then
\begin{equation}
  \frac{E_h - R_h}{h^q} \rightarrow \text{constant}
\end{equation}
and
\begin{equation}
  \frac{2^p - F_h}{h^{q-p}} \rightarrow \text{constant}
\end{equation}
which implies
\begin{equation}
  \log | 2^p - F_h| \approx \log|\text{constant}| + (q-p) \log h
\end{equation}
}

@@

\frame{\frametitle{When do we have an AEX?}

  Deriving an AEX is an exercise in Taylor expansions. If
  \begin{center}
    our functions are not many times differentiable
  \end{center}
  then the foundation crumbles.
  \vspace{1cm}
    \newline
  Every AEX refers to the exact value of the $A_h$. If
  $$\hat{A}_h \approx A_h$$
  is not a good approximation, then
  \begin{center}
    $\hat{A}_h$ will not behave nicely
  \end{center}
}

@@

\frame{\frametitle{Molecular dynamics: Truncation of force-fields}

  \begin{itemize}
  \item Simple simulation with 4 ions
  \item Each ion is attached to the origin with a spring
  \item The electrostatic forcefields are trunctated.
  \end{itemize}
}
